---
title: "The Typology of Causative-Applicative Syncretism"
author: "Bradley McDonnell"
date: "December 7, 2018 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
<style>
p, ul, ol, li, code {
  font-size: 20pt;
}

.glosstable table th {
  border: 0px;
  background-color: white;
  font-weight: normal;
}

.gr {
  font-variant: small-caps;
}
</style>



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(magrittr)
library(forcats)
library(leaflet) # for creating maps
library(wesanderson) # for custom colors
library(knitr) # for tables
library(kableExtra) # for tables

## Load datasets
wals_200_caus_appl <- read_tsv("../data/wals_200_caus_appl.csv")
wals_200_syncretism <- read_tsv("../data/wals_200_syncretism.csv")

## Define colors for maps
pal <- colorFactor(palette = wes_palette(name = "Zissou1")[c(2,5)],
                   domain = levels(wals_200_syncretism$causative_111a)
                   )
pal2 <- colorFactor(palette = wes_palette(name = "Zissou1")[c(5)],
                    domain = levels(wals_200_syncretism$reported)
                    )
```

```{r, load_refs, echo=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = "authoryear", 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
causApplBib <- ReadBib("Causative-Applicative.bib", check = FALSE)
```

---
# Goals

1. An overview of causatives and applicatives in the world's languages.
--

1. Worldwide distribution of causative-applicative syncretism.
--

  * Distribution within a subset of the WALS 200 sample.
--

1. Challenges & prospects creating a typology of causative-applicative syncretism.

---

class: inverse, center, middle

# Causatives

---

# WALS map for causatives `r Citep(causApplBib, "song2013nonperiphrastic")`

```{r out.width='100%', fig.height=6, eval=require('leaflet'), echo=FALSE}
causative_map <- wals_200_caus_appl %>%
  na.omit() %>%
  leaflet() %>%
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%

  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(causative_111a), 
                   weight = 1, 
                   fillColor = ~pal2(causative_111a),
                   fillOpacity = 0.75,
                   opacity = 1,
                   radius =  3, 
                   label=~as.character(name)
  ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~causative_111a,
            title = "Causatives",
            opacity = 1
  )
causative_map
```

---

# Causatives 

```{r, echo=FALSE}
wals_200_caus_appl %>%
  na.omit() %>%
  group_by(causative_111a) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html', 
        col.names = c("Causatives",
                      "Total (n)",
                      "Total (%)")
        )  %>%
  kable_styling()
```
---


class: inverse, center, middle

# Applicatives


---

# WALS map for applicatives `r Citep(causApplBib, "polinsky2013applicative")`

```{r, 'appl-map', out.width='100%', fig.height=6, eval=require('leaflet'), echo=FALSE}
applicative_map <- wals_200_caus_appl %>%
  na.omit() %>%
  leaflet() %>% 
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(applicative_109a), 
                   weight = 1, 
                   fillColor = ~pal2(applicative_109a),
                   fillOpacity = 0.75,
                   opacity = 1,
                   radius = 3, 
                   label=~as.character(name)
  ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~applicative_109a,
            title = "Applicatives",
            opacity = 1
  )
applicative_map
```

---

# Applicatives 

```{r, echo=FALSE}
wals_200_caus_appl %>%
  na.omit() %>%
  group_by(applicative_109a) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html',
        col.names = c("Applicatives",
                      "Total (n)",
                      "Total (%)")
        )  %>%
  kable_styling()
```

---

# Causative & Applicatives 

```{r, echo=FALSE}
caus_appl_table <- wals_200_caus_appl %>%
  na.omit() %>%
  group_by(causative_111a, applicative_109a) %>%
  summarize(n=n()) 

caus_appl_table %>%
  kable(format = 'html', 
        col.names = c("Causatives", 
                      "Applicatives", 
                      "Total (n)")
        )  %>%
  kable_styling()

caus_appl_table_df <- caus_appl_table %>%
  ungroup %>%
  as.data.frame()
```

---

# Causative & Applicatives

* Most langauges have a **causative** but *no* **applicative**
  * (n = `r caus_appl_table_df[3,3]`)

--
* Many languages have a **causative** *and* an **applicative** 
  * (n = `r caus_appl_table_df[4,3]`)

--
* A small number of langauges have *neither* a **causative** *nor* an **applicative**
  * (n = `r caus_appl_table_df[1,3]`)

--
* Very *few* langauges have an **applicative** but *no* **causative** 
  * (n = `r caus_appl_table_df[2,3]`)

---

class: inverse, center, middle

# Causative-applicative syncretism

---
# Causative-applicative syncretism map

```{r, 'sync-map', out.width='100%', fig.height=6, eval=require('leaflet'), echo=FALSE}
syncretism_map <- leaflet(wals_200_syncretism) %>%
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(syncretism), 
                   weight = 1, 
                   radius = 4, 
                   fillColor = ~pal(syncretism), 
                   fillOpacity = 0.75,
                   opacity = 1,
                   label=~as.character(name)
                   ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~syncretism,
            title = "Syncretism",
            opacity = 1
  )
syncretism_map
```

---
# Causative-applicative syncretism

```{r, echo=FALSE}
wals_200_syncretism %>%
  group_by(syncretism) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html', 
        col.names = c("Syncretism", 
                      "Total (n)",
                      "Total (%)")
        )  %>%
  kable_styling()
```

---

# References

```{r, 'refs', results='asis', echo=FALSE}
PrintBibliography(causApplBib, start = 1, end = 2)
```

---